---
title: "Olympic Medals"
author: "AndrewJ"
date: "6 August 2016"
output: html_document
---

# Description

As an exercise in increasing my `ggplot2` skills, replicate as best we can the plots of Olympic medals from chapter 3 of *Data Visualisation* by Andy Kirk. 

# Setup

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(cowplot)
library(grid)
library(RColorBrewer)
library(formattable)
theme_set(theme_minimal())
```

# Create the data

The data is augmented to 
```{r}
medals <- data_frame(
  Country = as.factor(c("United States of America", "People's Republic of China", 
              "Russian Federation", "Great Britain", "Australia", "Germany", 
              "France", "Republic of Korea", "All")),
  `2008` = c(110, 100, 72, 47, 46, 41, 40, 31, 951),
  `2004` = c(103, 63, 92, 30, 49, 49, 33, 30, 929),
  `2000` = c(92, 59, 88, 28, 58, 56, 38, 28, 925),
  `1996` = c(101, 50, 63, 15, 41, 65, 37, 27, 842),
  `1992` = c(108, 54, 112, 20, 27, 82, 29, 29, 815)
)

# Set up the comparisons for the last plot between China and Germany
medals$Compare <- as.factor(c(0, 1, 0, 0, 0, 2, 0, 0, 0))

medals_tidy <- medals %>%
  arrange(`2008`) %>%
  gather(Year, Medals, `2008`:`1992`) %>%
  filter(Year != "1992" | Country != "Russian Federation")

sorted_countries <-
  medals_tidy %>%
  filter(Year == "2008", Country != "All") %>%
  select(Country)
sorted_countries <- as.character(sorted_countries$Country)
```

# Plots

Floating bar chart of medal ranges over all the years (page 73)

```{r}
medals_min_max <- 
  medals_tidy %>%
  group_by(Country) %>%
  summarise(Min = min(Medals), Max = max(Medals))
```

```{r}
medals_min_max %>%
  filter(Country != "All") %>%
  ggplot() +
  geom_crossbar(
    aes(x = Country, y = Min, ymin = Min, ymax = Max),
    fill = "lightblue", width = 0.3, linetype = 0) +
  coord_flip() +
  scale_x_discrete(limits = rev(sorted_countries)) +
  scale_y_continuous(breaks = seq(0, 120, 20), limits = c(0, 120)) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major.y = element_blank())
```

Floating bar chart of percentages (page 74)

```{r}
# Add percentages of "All". Not sure how to do this in dplyr
medals_min_max$MinPct <- as.numeric(sapply(medals_min_max$Min, 
                                function (x) x/medals_min_max[1, "Min"]))
medals_min_max$MaxPct <- as.numeric(sapply(medals_min_max$Max, 
                                function (x) x/medals_min_max[1, "Max"]))

medals_min_max %>%
  filter(Country != "All") %>%
  mutate(
    MinPct = percent(MinPct, digits = 1), 
    MaxPct = percent(MaxPct, digits = 1)) %>%
  ggplot() +
  geom_crossbar(
    aes(x = Country, y = MinPct, ymin = MinPct, ymax = MaxPct),
    fill = "lightblue", width = 0.3, linetype = 0) +
  coord_flip() +
  scale_x_discrete(
    limits = rev(sorted_countries)) +
  scale_y_continuous(
    breaks = percent(seq(0, 0.14, 0.02), digits = 1), 
    limits = c(0, 0.14)) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major.y = element_blank())
```

Line graph of medals by years by country (page 75). Not yet able to put the y-axis on the RHS.

```{r}
medals_tidy %>%
  filter(Country != "All") %>%
  ggplot(aes(x = Year, y = Medals, colour = Country, group = Country)) +
  geom_line(size = 1.1) +
  scale_y_continuous(
    breaks = seq(0, 120, 20), limits = c(0, 120)) +
  scale_colour_discrete(
    limits=rev(sorted_countries)) +
  scale_colour_brewer(palette = "Set2") +
  theme(
    panel.grid.major.x = element_blank(),
    axis.title = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.position = "top")
#ggdraw(switch_axis_position(p, axis = 'y'))
```

The final line plot on page 77, sorta. Not sure how to do the subtitle.

```{r}
colour1 <- "#7799ee"
colour2 <- "orange"

medals_tidy %>%
  filter(Country != "All") %>%
  ggplot(aes(x = Year, y = Medals, colour = Country, group = Country)) +
  geom_line(aes(size = Compare, colour = Compare)) +

  guides(colour = FALSE, size = FALSE) + 
  scale_y_continuous(limits = c(0, 120)) +
  scale_size_manual(values = c(1, 1.2, 1.2)) +
  scale_colour_manual(values = c("#eeeeee", colour2, colour1)) +

  theme(
    panel.grid.major.x = element_blank(),
    axis.title = element_blank(),
    axis.line.x = element_line(colour = "#999999", size = 1.2),
    axis.text.x = element_text(face = "bold", margin = margin(t = -5, b = 5)),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold", size = 10, hjust = 0)) +
    ggtitle("The Contrasting Fortunes of German and Chinese Olympic Success") +

  # LHS
  annotate(geom = "text", x = 0.75, y = 82, size = 3,
           label = "Germany\n10.1%", colour = colour1) +
  annotate(geom = "text", x = 0.75, y = 54, size = 3,
           label = "China\n6.6%", colour = colour2) +
  
  # RHS
  annotate(geom = "text", x = 5.25, y = 41, size = 3,
           label = "4.3%", colour = colour1) +
  annotate(geom = "text", x = 5.25, y = 100, size = 3,
           label = "10.5%", colour = colour2)

```

